---
title: "Covid91 vaccine study Final2023"
author: "Diana flores"
output:
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

data from this study is taken from FinalData


$H_0$: the vaccines has no effect on people within a subgroup

$H_A$: the vaccine has an effect on people within a subgroup



## Methods

The statistical analysis i will be using to sort the data will be cat~cat
because we are putting the sub groups into category



the methods we will use to display the data is  bar chart, cross table, and chi squared test

## sub categorize male data
```{r}
maleData = filter (FinalData,sex == "M")
```


## Males


## Graphical Results
```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")

```


This graph shows use that from the all male group of people who contracted the ones who received a vaccine had less symptoms then those with the placebo while the all male group that didnt contract covid showed no difference between those who were vaccinated and un vaccinated.


## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=maleData)
rowPerc(table1)
colPerc(table1)
```

The data shows us that 69.57% of the unvaccinated male group experienced symptons while only 30.43% of the vaccinated males experienced symptoms

It also tells us that from the male placebo group 2.55% of them contracted covid while only 1.15% of the vaccinated group contracted covid

## Inferential Result


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p value for the chi squared test is .0000000000001098 which indicates a very significant relationship between the variables therefore disproving the null hypothesis $H_A$ giving strong evidence for the presence of a association between the variables

For the fisher exact test the pvalue is .00000000000004631 giving a strong basis for the observed association between the variables to unlikely have occurred by chance alone.

It also gives us a Odds ratio showing theres a  2.242522% higher chance of the events odds happening in one group giving more evidence to the positive association of the variables and rejection of the null $H_A$

## sub categorizing female data

```{r}
femaleData = filter (FinalData,sex == "F")
```


## Females

## Graphical Results

```{r}
barchartGC(~infected + treatment, data=femaleData)
barchartGC(~infected+treatment, data=femaleData, type="percent")

```

This graph shows use that from the all female group of people who contracted covid-19 the ones who received a vaccine had less symptoms overall then those with the placebo. while the all female group that didnt contract covid showed almost no difference between those who were vaccinated and un vaccinated.


## Numerical Results

```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```
The data from this table shows that from the female group that were infected with covid-19 of those who were unvaxed 60% of them experienced symptoms while only 40% of people in the vaccinated group experienced symptoms

This data also depicts that within the female placebo group 2.04% were infected with covid but only 1.37% of the vaccinated group




## Inferential Result

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```


The data suggests a significant relationship between the treatment type (placebo vs. vaccine) and infection status based on Pearson's Chi-squared test (p-value = 0.0003). Therefore, we reject the null hypothesis $H_A$ , indicating that there is an association between the variables.


Fisher's exact test for count data on table2 reveals a odds ratio of (1.493076) with a 95% confidence interval of 1.199509 to 1.862988 suggesting a statistically significant positive association between the two categorical variables, indicating that the event's odds are about 1.49 times higher in one group compared to the other.

The p-value from fishers exact test (0.0002162) indicates strong evidence against the null hypothesis, suggesting a statistically significant association between the two categorical variables again disproving the null $H_A$

## subsetting lgtbq data

```{r}
LGBTQData = filter (FinalData,LGBTQ == "gay")
```
## LGBTQ
## Graphical Results

```{r}
barchartGC(~infected+treatment, data=LGBTQData)
barchartGC(~infected+treatment, data=LGBTQData, type="percent")

```

This Graph shows that within the lgbtq subgroup Those afflicted with covid 19 and that were vaccinated experienced more symptoms then their un vaccinated counter part. for the people who never contracted covid vaxxed or un-vaxxed it didn't show any difference


## Numerical Results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```

This table shows that within the all gay placebo group  15.38% experienced symptoms while 84.62% of the vaccinated group had symptoms

From the all gay placebo group only .66% of the unvaccinated got infected with covid-19 while a large 3.54% of vaccinated gay people got infected with covid_19


## Inferential Result
```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The p value within the LGBTQ group is  0.0000006668  showing that theres a significant association between the variables  so i am rejecting the null hyothesis $H_A$ and conclude the covid vaccine has an effect within the LGBTQ population

the fishers exact tesst p value is  1.029e-08 giving evidence for the extremely low probability of obtaining the observed results by chance alone, suggesting a highly significant association between the variables therefore disproving the null hypothesis $H_A$

The odds ratio of the fisher exact test is  0.181787 indicating that the event's odds are approximately 0.18 times lower in one variable group compared to the other, implying a significant difference between the two groups. Again adding to the body of evidence against the null hyopthesis $H_A$




## subset druggie data
```{r}
DrugUserData = filter(FinalData, DrugUser == "yes")
```
## Druggies
## Graphical Results

```{r}
barchartGC(~infected + treatment, data=DrugUserData)
barchartGC(~infected+treatment, data=DrugUserData, type="percent")

```

This graph shows that for druggies who were infected with covid-19 those who were vaccinated actually experienced more symptoms then those that were un-vaccinated. The druggies who didnt get infected with covid-19 wether vaccinated or un-vaccinated showed an insignificant difference


## Numerical Results

```{r}
table4 <- xtabs(~infected + treatment, data=DrugUserData)
rowPerc(table4)
colPerc(table4)
```

The table shows that for the placebo druggie group 11.67% of them experienced covid19 symptoms when they were infected with covid-19 while 88.33% of vaccinated drug users experienced symptons

The bottom table is showing that .4% of the placebo group got infected with  covid while 3.06% of the vaccinated group got infected with covid


## Inferential Result

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The p-value of .000000002842 from the Pearson's Chi-squared test indicates a highly significant association between the variables, leading to the rejection of the null hypothesis $H_A$.

Fisher's exact test yielded a highly significant p-value (2.32e-10) and a 95% confidence interval (0.0482 to 0.2791), indicating a strong association between the categorical variables, with the estimated odds ratio being 0.126 it suggest that the odds of the event occurring are substantially lower in one group compared to the other, supported by a 95 percent confidence interval ranging from 0.04819608 to 0.27914768 , this body of evidence leads me to reject the null hypothesis $H_A$ .

# Overall Results and Conclusions

In conclusion, the effectiveness of the COVID vaccine varies between subgroups. Staring with the all male group, we are rejecting the null hypothesis. Based on the bar chart, numerical graph, and the chisq and fisher tests, we can conclude that men who received the COVID vaccine and contracted it ultimately experience less symptoms than those who received the placebo. The men that did not get vaccinated and did not contract COVID showed little to no difference between eachother. In the group that did receive the placebo had an infection rate of 2.55%, and those of them who got the vaccine had an infection rate of 1.15%, showing that the vaccine is effective in men, therefore we are rejecting the null hypothesis for the male group. 

Moving onto the all female group, we are also rejecting the null hypothesis. Based on all of the tests we have run, we can conclude that women who received the COVID vaccine and contracted it also experienced less symptoms than the women who received the placebo. The female group that did not get vaccinated and did not contract COVID also showed little to no difference compared to each other. Within the group that did receive the placebo, they had an infection rate of 2.04%. The group that received the vaccine had an infection rate of 1.37%, concluding that the vaccine is effective in women, therefore rejecting the null hypothesis for the female subgroup.

Comparing the male and female subgroups, women had a smaller percentage of infection with the COVID vaccine and placebo than then men by 0.22% and 0.51%. It is not a big difference, but so far within the data analysis women experience less symptoms with the COVID vaccine. 

In the LGBTQ subgroup, we are rejecting the null hypothesis. Based on our tests, we can conclude that the COVID vaccine is effective in this subgroup, but negatively. We found out that the LGBTQ people who received the COVID vaccine and contracted it experienced significant symptoms than those who received the placebo. The LGBTQ group that did not get vaccinated and did not contract COVID were almost exactly the same, showing no difference. In the LGBTQ group that did receive the placebo had an infection rate of 0.66% and those who did receive the vaccine had an infection rate of 3.54%, a very significant difference between the two, therefore rejecting the null hypothesis since the vaccine had an effect on the group.

Finally, for the drug user group, they received the same result as the LGBTQ subgroup. We are rejecting the null hypothesis due to the drug users experiencing significantly more symptoms than the placebo group, showing that the COVID vaccine did have an effect, but not a positive one In the group of drug users that were vaccinated and contracted COVID, they had a 3.06% infection rate. The placebo group had a 0.4% infection rate showing a very big difference between the two and how much symptoms each group got, once again, we are rejecting the null.

Comparing all of the tests from most positive COVID affects to the more negative COVID affects, the female had the best results with the COVID vaccine, while the LGBTQ subgroup had it the worst, experiencing the most symptoms while having COVID even with the vaccine. Males had almost the same results as the females, just a little less and drug users had fewer symptoms by only 0.48%. I would recommend that only males and females that are not drug users or LGBTQ to get the COVID vaccine since the effect was more positive than it was for the other two subgroups.  